---
output:
  pdf_document: default
  html_document: default
---
<h1 id="0"></h1>

---
title: "Online sport gambling Descriptive analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load myData, include=FALSE}
library(readr)
library(dplyr)
library(plotly)
library(tidyr)
library(rmarkdown)
library(markdown)
library(knitr)
library(tinytex)
basetable <- read_csv("C:/Users/schauhan/Downloads/Final R Group 12/basetable.csv",
                      col_types = cols(RegDate = col_date(format = "%Y-%m-%d")))
```


####GROUP MEMBER: SIDDHANT CHAUHAN / SHIVAM SARIN / YEN CHUN, LIU


<h1 id="Introduction">Introduction</h1>

The goal of this report is to provide overall picture on the customers and insights. For this analysis report used four raw datasets plus information provided in the [codebook](https://iesegnet-my.sharepoint.com/:b:/g/personal/yenchun_liu_ieseg_fr/EWqOitNVhrJIj3aZHM08tIcBQDhdNKcSbmlHtTILgMLZnw?e=ba6qV5) to generate basetable. Charts are mainly analysis with RFM(Recency, Frequency, Monetary) and others like location etc. The Chart list below are some demonstrate of possible overview analysis using basetable through [R shiny](https://shivamsarin.shinyapps.io/R_Grp_12/). R shinny provides flexibility for users to choose from five variables(Gender, Country, Product ID, Application ID, User ID, Language).


<h2 id="Chart list">Chart list</h2>

  [1. Unique Users per Products](#1)

  [2. Frequency Plot for Poker Chips by months](#2)

  [3. Frequency Plot for User Aggregation by months](#3)

  [4. Frequency Plot - Application used](#4)

  [5. Frequency Plot - Age](#5)

  [6. Unique Users per Languages](#6)

  [7. Pie Chart of Churn](#7)

  [8. Unique Users per Country](#8)

  [9. Frequency Plot - Length of relationship (Days)](#9)

  [10.Unique Users per Gender](#10)

  [11.Average number of days between betting sessions - Poker Chips](#11)

  [12.Average number of days between betting sessions - User Aggregation](#12)

  [13.Product wise - Profit vs Loss](#13)

  [14.Product wise - Profit% vs Loss%](#14)

  [15.Winnings Vs Profit Vs Stakes](#15)

  [16.Freq vs Profit for User Aggregation](#16)

  [17.Stakes Vs Profit](#17)



<h3 id="1">1. Unique Users per Products</h3>

```{r Unique Users per Products, echo=FALSE, cache=TRUE}
# temp <- subset(basetable, is.na(basetable$ProductID))
product_wise_players <- basetable %>%
  group_by(ProductID) %>%
  summarise(unique_players = n_distinct(USERID))
product_wise_players <- product_wise_players[complete.cases(product_wise_players),]

product_wise_players_PC <- unique(basetable[,c("USERID","PC_Recency")])
product_wise_players_PC <- sum(!is.na(product_wise_players_PC$PC_Recency))

product_wise_players <- rbind(product_wise_players,c("Poker_Chips",product_wise_players_PC))
product_wise_players$unique_players <- as.numeric(product_wise_players$unique_players)

product_wise_players %>%
  group_by(ProductID) %>%
  summarize(unique_players = unique_players) %>%
  plot_ly(labels = ~ProductID, values = ~unique_players) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "",  showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

rm(product_wise_players_PC)
rm(product_wise_players)

```
[Back to list](#Chart list)


<h3 id="2">2. Frequency Plot for Poker Chips by months</h3>

```{r Frequency Plot for Poker Chips by months, echo=FALSE, cache=TRUE}
############ Frequency Plot for Poker Chips by months ###########

freq_by_months <- unique(basetable[,c("USERID","PC_Freq_sessions_first3Months",
                                      "PC_Freq_sessions_mid3Months",
                                      "PC_Freq_sessions_last2Months")])
names(freq_by_months) <- c("USERID","First 3 Months",
                             "Mid 3 Months",
                             "Last 2 Months")
freq_by_months <- gather(freq_by_months,Freq,value,-USERID)
freq_by_months <- freq_by_months[complete.cases(freq_by_months),]
freq_by_months$value <- as.character(freq_by_months$value)

x = factor(freq_by_months$Freq,c("USERID","First 3 Months",
                                 "Mid 3 Months",
                                 "Last 2 Months"))
y = freq_by_months$value

plot_ly(y=y, x=x, histfunc='sum', type = "histogram") %>%
  layout(title = "",
         yaxis=list(type='linear'),
         xaxis=list(tickangle = 315))
```
[Back to list](#Chart list)

<h3 id="3">3. Frequency Plot for User Aggregation by months</h3>
```{r Frequency Plot for User Aggregation by months, echo=FALSE, cache=TRUE}
############ Frequency Plot for User Aggregation by months ###########

freq_by_months <- unique(basetable[,c("USERID","UA_Freq_bets_first3Months",
                                      "UA_Freq_bets_mid3Months",
                                      "UA_Freq_bets_last2Months")])
names(freq_by_months) <- c("USERID","First 3 Months",
                           "Mid 3 Months",
                           "Last 2 Months")
freq_by_months <- gather(freq_by_months,Freq,value,-USERID)
freq_by_months <- freq_by_months[complete.cases(freq_by_months),]
freq_by_months$value <- as.character(freq_by_months$value)

x = factor(freq_by_months$Freq,c("USERID","First 3 Months",
                                 "Mid 3 Months",
                                 "Last 2 Months"))
y = freq_by_months$value

plot_ly(y=y, x=x, histfunc='sum', type = "histogram") %>%
  layout(title = "",
         yaxis=list(type='linear'),
         xaxis=list(tickangle = 315))

rm(freq_by_months)
rm(x)
rm(y)
```
[Back to list](#Chart list)


<h3 id="4">4. Frequency Plot - Application used</h3>
```{r Frequency Plot - Application used, echo=FALSE, cache=TRUE}
############ Frequency Plot for Application used ###########

freq_by_application <- unique(basetable[,c("USERID","ApplicationID")])
freq_by_application <- freq_by_application[complete.cases(freq_by_application),]

freq_for_ordered_names <- freq_by_application %>%
  group_by(ApplicationID) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq))

freq_by_application$value <- as.character(1)

x = factor(freq_by_application$ApplicationID, levels = freq_for_ordered_names$ApplicationID)
y = freq_by_application$value

plot_ly(y=y, x=x, histfunc='sum', type = "histogram") %>%
  layout(title = "",
         yaxis=list(type='linear'),
         xaxis=list(tickangle = 315))

rm(freq_by_application)
rm(freq_for_ordered_names)
rm(x)
rm(y)

```
[Back to list](#Chart list)


<h3 id="5">5. Frequency Plot - Age</h3>
```{r Frequency Plot - Age, echo=FALSE, cache=TRUE}
############ Frequency Plot for Age ###########

freq_by_age <- unique(basetable[,c("USERID","AGE")])
freq_by_age <- freq_by_age[complete.cases(freq_by_age),]

x = freq_by_age$AGE

plot_ly(x=x, type = "histogram") %>%
  layout(title = "")

rm(freq_by_age)
rm(x)

```
[Back to list](#Chart list)


<h3 id="6">6. Unique Users per Languages</h3>
```{r Unique Users per Languages, echo=FALSE, cache=TRUE}
#################### Pie chart Languages #################

language_wise_players <- basetable %>%
  group_by(Language) %>%
  summarise(unique_players = n_distinct(USERID))

language_wise_players <- language_wise_players[complete.cases(language_wise_players),]

language_wise_players %>%
  group_by(Language) %>%
  summarize(unique_players = unique_players) %>%
  plot_ly(labels = ~Language, values = ~unique_players) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "",  showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

rm(language_wise_players)
```
[Back to list](#Chart list)


<h3 id="7">7. Pie Chart of Churn</h3>
```{r Pie chart Churn, echo=FALSE, cache=TRUE}
#################### Pie chart Churn #################

Churn_wise_players <- basetable %>%
  group_by(Churn) %>%
  summarise(players = n())

Churn_wise_players <- Churn_wise_players[complete.cases(Churn_wise_players),]

Churn_wise_players %>%
  group_by(Churn) %>%
  summarize(players = players) %>%
  plot_ly(labels = ~Churn, values = ~players) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "",  showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

rm(Churn_wise_players)
```
[Back to list](#Chart list)


<h3 id="8">8. Unique Users per Country</h3>
If it does not appear please use browser to view.
```{r Unique Users per Country, echo=FALSE, cache=TRUE}
#################### Map Country ####################


freq_by_country <- unique(basetable[,c("USERID","Country")])
freq_by_country <- freq_by_country[complete.cases(freq_by_country),]

freq_by_country_names <- freq_by_country %>%
  group_by(Country) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq))

df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')

freq_by_country_names <- merge(x=freq_by_country_names, y= df,by.x = c("Country"),
                         by.y = c("COUNTRY"), all.x = T)

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

plot_geo(freq_by_country_names) %>%
  add_trace(
    z = ~freq, color = ~freq, colors = 'Blues',
    text = ~Country, locations = ~CODE, marker = list(line = l)
  ) %>%
  colorbar(title = 'Frequency') %>%
  layout(
    title = '',
    geo = g
  )
rm(list=setdiff(ls(), "basetable"))
```
[Back to list](#Chart list)


<h3 id="9">9. Frequency Plot - Length of relationship (Days)</h3>
```{r Frequency Plot for Age, echo=FALSE, cache=TRUE}
############ Frequency Plot for Age ###########

freq_by_lor <- unique(basetable[,c("USERID","UA_Lor_firstPlay")])
freq_by_lor <- freq_by_lor[complete.cases(freq_by_lor),]

x = freq_by_lor$UA_Lor_firstPlay

plot_ly(x=x, type = "histogram") %>%
  layout(title = "")

rm(freq_by_lor)
rm(x)
```
[Back to list](#Chart list)


<h3 id="10">10. Unique Users per Gender</h3>
```{r Pie chart Gender, echo=FALSE, cache=TRUE}
#################### Pie chart Gender #################

Gender_wise_players <- basetable %>%
  group_by(Gender) %>%
  summarise(players = n_distinct(USERID))

Gender_wise_players <- Gender_wise_players[complete.cases(Gender_wise_players),]

Gender_wise_players %>%
  group_by(Gender) %>%
  summarize(players = players) %>%
  plot_ly(labels = ~Gender, values = ~players) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "",  showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

rm(Gender_wise_players)
```
[Back to list](#Chart list)


<h3 id="11">11. Average number of days between betting sessions - Poker Chips</h3>
```{r Frequency Plot for Poker Chips - IPT, echo=FALSE, cache=TRUE}
############ Frequency Plot for Poker Chips - IPT ###########

freq_by_IPT <- unique(basetable[,c("USERID","PC_IPT_mean")])
library(pracma)
freq_by_IPT$PC_IPT_mean <- ceil(freq_by_IPT$PC_IPT_mean)
freq_by_IPT <- freq_by_IPT[complete.cases(freq_by_IPT),]

x = freq_by_IPT$PC_IPT_mean

plot_ly(x=x, type = "histogram") %>%
  layout(title = "")

rm(freq_by_IPT)
rm(x)
```
[Back to list](#Chart list)


<h3 id="12">12. Average number of days between betting sessions - User Aggregation</h3>
```{r Frequency Plot for User Aggregation - IPT, echo=FALSE, cache=TRUE}
############ Frequency Plot for User Aggregation - IPT ###########

freq_by_IPT <- unique(basetable[,c("USERID","UA_IPT_mean")])
freq_by_IPT$UA_IPT_mean <- ceil(freq_by_IPT$UA_IPT_mean)
freq_by_IPT <- freq_by_IPT[complete.cases(freq_by_IPT),]

x = freq_by_IPT$UA_IPT_mean

plot_ly(x=x, type = "histogram") %>%
  layout(title = "")

rm(freq_by_IPT)
rm(x)
```
[Back to list](#Chart list)


<h3 id="13">13. Product wise - Profit vs Loss</h3>
```{r Product wise - Profit vs Loss, echo=FALSE, cache=TRUE}
######Product wise - Profit vs Loss#######
profit_loss <- unique(basetable[,c("USERID","ProductID","UA_Mon_profit")])
profit_loss$Mon <- ifelse(profit_loss$UA_Mon_profit > 0,"Profit","Loss")
profit_loss$UA_Mon_profit <- abs(profit_loss$UA_Mon_profit)
profit_loss <- profit_loss[complete.cases(profit_loss),]

profit_loss1 <- profit_loss %>%
  group_by(ProductID) %>%
  summarise(Profit = sum(UA_Mon_profit[Mon=="Profit"],na.rm = T),
            Loss = sum(UA_Mon_profit[Mon=="Loss"],na.rm = T),
            Profit_Percentage = Profit*100/(Profit+Loss),
            Loss_Percentage = Loss*100/(Profit+Loss))

profit_loss <- unique(basetable[,c("USERID","PC_Mon_profit")])
profit_loss$Mon <- ifelse(profit_loss$PC_Mon_profit > 0,"Profit","Loss")
profit_loss$PC_Mon_profit <- abs(profit_loss$PC_Mon_profit)

profit_loss2 <- profit_loss %>%
  summarise(Profit = sum(PC_Mon_profit[Mon=="Profit"],na.rm = T),
            Loss = sum(PC_Mon_profit[Mon=="Loss"],na.rm = T),
            Profit_Percentage = Profit*100/(Profit+Loss),
            Loss_Percentage = Loss*100/(Profit+Loss))
profit_loss2$ProductID <- "Poker Chips"
profit_loss <- rbind(profit_loss1,profit_loss2)

rm(profit_loss1)
rm(profit_loss2)

profit_loss <- profit_loss %>%
  group_by(ProductID) %>%
  arrange(desc(Profit))

x = factor(profit_loss$ProductID, levels = profit_loss$ProductID)
y1 = profit_loss$Loss
y2 = profit_loss$Profit

plot_ly(x = ~x, y = ~y1, type = 'bar', name = 'Loss') %>%
  add_trace(y = ~y2, name = 'Profit') %>%
  layout(title = '',
    yaxis = list(title = 'Value'), barmode = 'stack',
         xaxis=list(title = 'Products',tickangle = 315))
```
[Back to list](#Chart list)


<h3 id="14">14. Product wise - Profit% vs Loss%</h3>
```{r Product wise - Profit% vs Loss%, echo=FALSE, cache=TRUE}
profit_loss <- profit_loss %>%
  group_by(ProductID) %>%
  arrange(desc(Profit_Percentage))

x = factor(profit_loss$ProductID, levels = profit_loss$ProductID)

y1 = profit_loss$Loss_Percentage
y2 = profit_loss$Profit_Percentage

plot_ly(x = ~x, y = ~y1, type = 'bar', name = 'Loss %') %>%
  add_trace(y = ~y2, name = 'Profit %') %>%
  layout(title = '',
          yaxis = list(title = 'Value %'), barmode = 'stack',
         xaxis=list(title = 'Products',tickangle = 315))
```
[Back to list](#Chart list)

<h3 id="15">15. Winnings Vs Profit Vs Stakes</h3>
```{r Winnings Vs Profit Vs Stakes, echo=FALSE, cache=TRUE}
######### Winnings Vs Profit Vs Stakes ##########
plot_ly(basetable, x =1:nrow(basetable), y = basetable$UA_rMon_profit, name = 'Profit',type = 'scatter', mode = 'lines')%>%
  add_trace(y = basetable$UA_rMon_wins, name = ' Wins', mode = 'lines') %>%
  add_trace(y = basetable$UA_rMon_Stakes, name = 'Stakes', mode = 'lines')  %>%
  layout(title="")

```
[Back to list](#Chart list)

<h3 id="16">16. Freq vs Profit for User Aggregation</h3>
```{r Freq vs Profit for User Aggregation, echo=FALSE, warning=FALSE, cache=TRUE}
##############Bets Vs Profits/ User ID for UA ######################
plot_ly(basetable, x = basetable$UA_rFreq_bets, y =basetable$UA_rMon_profit, text = basetable$USERID, type = 'scatter', mode = 'markers')%>%
  layout(title="")
```
[Back to list](#Chart list)


<h3 id="17">17. Stakes Vs Profit</h3>
```{r Stakes Vs Profit, echo=FALSE, cache=TRUE}
############# Poker Profit vs Freq vs Buy######################3
plot_ly(basetable, x =1:nrow(basetable), y =basetable$PC_rMon_profit, name = 'Profit',type = 'scatter', mode = 'lines')%>%
    add_trace(y = basetable$PC_rMon_Buy, name = 'Stakes', mode = 'lines')  %>%
  layout(title="")
rm(basetable)
rm(x)
```

[Back to list](#Chart list)

[Back to top](#0)
